---
title: "2015 State competition, Question 18"
author: "Fred Boehm"
date: "11/08/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem statement 
The quadratic equation $$x^2+ px + 2p = 0$$ has solutions $x = a$ and $x = b$. If the

quadratic equation $$x^2+ cx + d = 0$$

 has solutions $x = a + 2$ and $x = b + 2$, what is the value of $d$?


## Approach

We'll rely on factorization of quadratics to solve this. 

First, note that we're given the two solutions to the first equation, $a$ and $b$. This tells us that:

$$(x - a)(x-b) = 0 = x^2+ px + 2p$$
We can expand the left-hand side to get:

$$(x-a)(x-b) = x^2 - (a + b)x + ab = x^2 + px + 2p$$
For the two sides of the equation to be equal, their coefficients for the $x$ terms must be equal. 

That is, we must have $$-(a + b) = p$$ and $$ab = 2p$$

From these two equations, we see that:

$$-2(a + b) = 2p = ab$$

Using "guess and check", notice that $a = 2$ and $b = -1$ is one solution to the above equation.


We're also given the two solutions to the second equation, $a + 2$ and $b + 2$. 

$$(x - (a + 2))(x - (b + 2)) = 0 = x^2+ cx + d$$
Using the same approach, ie, matching the coefficients of the powers of $x$, we see that

$$(a + 2)(b + 2) = d$$

Plug in the values of $a$ and $b$ that solved the first quadratic equation to get:

$$d = (2 + 2)(-1 + 2) = 4$$




## Resources







## Git info

```{r} 
git2r::last_commit()
```


